---
title: "Chap18 - Exercise"
author: "Me"
date: "10/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#2 
For oil data
estimate variable importance scores. Does large number of classes affect process of qualifying importance


```{r}
library(caret)
```

Get matrix correlation and p values
```{r}
cor.prob = function(X, dfr=nrow(X)-2){
    R = cor(X)
    above = row(R) = col(R)
    r2 = R[above]^2
    Fstat = r2 * dfr/(1 - r2)
    R[above] = 1 - pf(Fstat, 1, dfr)
    R[row(R) == col(R)] = NA
    R
}
```

Extract largest cor
```{r}
largest_cors = function(X){
    pred_cors = cor( X )

    # Convert the matrix above into a dataframe with an entry for each element of the correlation matrix:
    cor_DF = data.frame( first_variable=rep(rownames(pred_cors),ncol(pred_cors)),
                         second_variable=rep(colnames(pred_cors),each=nrow(pred_cors)),
                         cor=as.vector(pred_cors) )

    upper_mask = as.vector(upper.tri(pred_cors)) # which elements are on the upper triangular (to avoid duplicates)

    # Order these pairwise-correlation by their absolute correlation:
    #
    pairwise_cor = cor_DF[upper_mask,]
    pairwise_cor$abs_cor = abs(pairwise_cor$cor)
    pairwise_cor = pairwise_cor[ order( pairwise_cor$abs_cor, decreasing=TRUE ), ]
    pairwise_cor$abs_cor = NULL
    pairwise_cor
}
```


```{r}
library(AppliedPredictiveModeling)
library(C50)
library(CORElearn)
```

```{r}
data(oil)
filterVarImp(fattyAcids, oilType)
```

try second method
```{r}
DF = cbind(fattyAcids, oilType)

relief_values = attrEval(oilType ~., data = DF, estimator = 'ReliefFequalK', ReliefIterations = 50)
```
#3 
```{r}
data(abalone)
str(abalone)
```

Data consists of measurements of type(male, female, infant), longest shell measurement, the diameter, height, several weights. Outcome is the number of rings.
```{r}
head(abalone)
```

a - plot the data to assess relationships between predictors and outcome
b - use scatterplots and correlation to understand how predictors relate to one another
c - estimate variable importance. Develop an approach to determine reduced set of nonredundant predictors
d - apply PCa to continous to determine how many distinct info are in the data

```{r}
cn = colnames(abalone)

for(ci in 2:8){
  plot(abalone[,ci], abalone$Rings, main = cn[ci])
  cat('Press enter to continue')
  line = readline()
}
```

#Save scatterplots that look interesting

Height vs rings
```{r}
par(mfrow = c(1,2))
plot(Rings ~ Height, data = abalone)
orig_limits = par("usr")
m_orig = lm( Rings ~ Height, data=abalone )
abline(m_orig)
grid()
```


```{r}
largest_cors(abalone[,-1])
```

```{r}
findCorrelation(cor(abalone[,-1]), cutoff = .75)
```
Assess importance
```{r}
VI = filterVarImp( abalone[,1:(dim(abalone)[2]-1)], abalone[,dim(abalone)[2]] )
```
Try another method
```{r}
reliefvalues = attrEval( Rings ~ ., data=abalone[,-1], estimator="RReliefFequalK" )
```
Apply PCA to continuous data
```{r}
pca_out = prcomp( abalone[,-1], center=T, scale=T )
```

















