---
title: "Chap5 - Peformance in regression models"
author: "Me"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To compute for model performance, the observed and predicted outcomes should be stored in vectors. FOr regression, vectors should be numeric. Here's an example of manually creating a vector
```{r}
observed = c(.22, .83, -.12, .89, -.23, -1.3, -.15, -1.4, 
             .62, .99, -.18, -.32, .34, -.3, .04, -.87,
             .55, -1.3, -1.15,.2)

predicted = c(.24, .78, -.66, .53, .7, -.75, -.41, -.43, 
             .49, .79, -1.19, -.06, .75, -.07, .43, -.42,
             -.25, -.64, -1.26, -0.7)

resid = observed - predicted
summary(resid)
```

An important first step in evaluating the quality of the model is to visualize the results. First, a plot of the observed values against the predicted helps one to understand how the model fits. Also a plot of residuals vs predicted values can help uncover patterns in the model predictions. 

```{r}
#observed vs predicted
library(ggplot2)

ggplot(data = NULL, aes(observed, predicted))+
  geom_point()+
  geom_abline(slope = 1, intercept = 0)
```
Plot the predicted values vs residuals
```{r}
plot(predicted, resid)
abline(h = 0, col = 'darkblue')
```
The caret package contains function for calculating RMSe and R2

```{r}
library(caret)
R2(predicted, observed)
RMSE(predicted, observed)
```
There are different formulas for R2. The default R2 function uses the square correlation coefficient. Base R contains function to compute correlation including spearman's rank correlation
```{r}
#simple cor
cor(predicted, observed)

#rank correlation
cor(predicted, observed, method = 'spearman')
```


























