---
title: "Chap13 - Exercises"
author: "Me"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 
Use hepatic injury data from previous exercise set. 
a - work with the same training and testing and preprocessing. Using same classification statistic, build non linear models. Which models has the best predictive ability? Does non linear structure help improve model classification?
b - what are the top 5 important predictors
c-  combine biological and chemical to one predictor set. Which model has the best performance? What are the top 5?
d - which model would you recommend

```{r}
library(caret)
library(AppliedPredictiveModeling)
library(pROC)
```
```{r}
data(hepatic)
library(forcats)
any_damage = fct_collapse(injury, 'Yes' = c('Mild','Severe'), No = 'None')
table(any_damage)
```
## Biological predictors
```{r}
zv_cols = nearZeroVar(bio)
X_bio = bio[,-zv_cols]

findLinearCombos(X_bio)
```
No linear combination of predictors

BUILD MODELS
```{r}
ctrl =  trainControl( summaryFunction=twoClassSummary, classProbs=TRUE )
```

##MDA
```{r}
set.seed(100)
mda_model = train(X_bio, any_damage, method = 'mda', tuneGrid = expand.grid(subclasses = 1:5), metric = 'ROC', trControl = ctrl)
mda_preds = predict(mda_model, X_bio, type = 'prob')
mda_roc = roc(response = any_damage, predictor = mda_preds[,1])
mda_auc = mda_roc$auc[1]
mda = list(model = mda_model, roc = mda_roc, auc = mda_auc)
mda
```

##NN
```{r}
set.seed(100)
nnet_grid = expand.grid(size = 1:5, decay = c(0, .1, 1, 2))
nnet_model = train(X_bio, any_damage, method = 'nnet', preProcess = c('center','scale','spatialSign'),
                   tuneGrid = nnet_grid, metric = 'ROC', trace = F, maxit = 2000, trControl = ctrl)
nnet_preds = predict(nnet_model, X_bio, type = 'prob')
nnet_roc = roc(response = any_damage, predictor = nnet_preds[,1])
nnet_auc = nnet_roc$auc[1]
nnet = list(model = nnet_model, roc = nnet_roc, auc = nnet_auc)
nnet
```
##SVm
```{r}
set.seed(100)
sigma_est = kernlab::sigest(as.matrix(X_bio))
svm_grid = expand.grid(sigma = sigma_est[1], C= 2^seq(-4,4))
svm_model = train(X_bio, any_damage, method = 'svmRadial', tuneGrid = svm_grid, preProcess = c('center','scale'), metric = 'ROC', fit = F, trControl = ctrl)
svm_preds = predict(svm_model, X_bio, type = 'prob')
svm_roc = roc(response = any_damage, predictor = svm_preds[,1])
svm_auc = svm_roc$auc[1]
svm = list(model = svm_model, roc = svm_roc, auc = svm_auc)
```

#KNN
```{r}
set.seed(100)
knn_model = train(X_bio, any_damage, method = 'knn', tuneLength = 20, preProcess = c("center",'scale'), metric = 'ROC', trControl = ctrl)
knn_preds = predict(knn_model, X_bio, type ='prob')
knn_roc = roc(response = any_damage, predictor = knn_preds[,1])
knn_auc = knn_roc$auc[1]
knn = list(model = knn_model, roc = knn_roc, auc = knn_auc)
```
Naive bayes
```{r}
set.seed(100)
nb_model = train(X_bio, any_damage, method = 'nb', metric = 'ROC', trControl = ctrl)
nb_preds = predict(nb_model, X_bio, type = 'prob')
nb_roc = roc(response = any_damage, predictor = nb_preds[,1])
nb_auc = nb_roc$auc[1]
nb = list(model = nb_model, roc = nb_roc, auc = nb_auc)
```

```{r}
bio_result = list( mda = mda, nnet=nnet, svm=svm, knn=knn, nb=nb )

df = rbind(data.frame(name = 'MDA', auc = bio_result$mda$auc),
           data.frame(name = 'NNET', auc = bio_result$nnet$auc),
           data.frame(name = 'SVM', auc = bio_result$svm$auc),
           data.frame(name = 'KNN', auc = bio_result$knn$auc),
           data.frame(name = 'NB', auc = bio_result$nb$auc))
df
```

```{r}
varImp(bio_result$mda$model)
```

## Chemical predictors
```{r}
zv_cols = nearZeroVar(chem)
X_chem = chem[,-zv_cols]

flc = findLinearCombos(X_chem)
X_chem = X_chem[, -flc$remove]
```

#Build models

```{r}
set.seed(100)
mda_model = train(X_chem, any_damage, method = 'mda', tuneGrid = expand.grid(subclasses = 1:5), metric = 'ROC', trControl = ctrl)
mda_preds = predict(mda_model, X_chem, type = 'prob')
mda_roc = roc(response = any_damage, predictor = mda_preds[,1])
mda_auc = mda_roc$auc[1]
mda = list(model = mda_model, roc = mda_roc, auc = mda_auc)
```

```{r}
set.seed(100)
nnet_grid = expand.grid(size = 1:5, decay = c(0, .1, 1, 2))
nnet_model = train(X_chem, any_damage, method = 'nnet', preProcess = c('center','scale','spatialSign'),
                   tuneGrid = nnet_grid, metric = 'ROC', trace = F, maxit = 2000, trControl = ctrl)
nnet_preds = predict(nnet_model, X_chem, type = 'prob')
nnet_roc = roc(response = any_damage, predictor = nnet_preds[,1])
nnet_auc = nnet_roc$auc[1]
nnet = list(model = nnet_model, roc = nnet_roc, auc = nnet_auc)
```

```{r}
set.seed(100)
sigma_est = kernlab::sigest(as.matrix(X_chem))
svm_grid = expand.grid(sigma = sigma_est[1], C= 2^seq(-4,4))
svm_model = train(X_chem, any_damage, method = 'svmRadial', tuneGrid = svm_grid, preProcess = c('center','scale'), metric = 'ROC', fit = F, trControl = ctrl)
svm_preds = predict(svm_model, X_chem, type = 'prob')
svm_roc = roc(response = any_damage, predictor = svm_preds[,1])
svm_auc = svm_roc$auc[1]
svm = list(model = svm_model, roc = svm_roc, auc = svm_auc)
```

```{r}
set.seed(100)
knn_model = train(X_chem, any_damage, method = 'knn', tuneLength = 20, preProcess = c("center",'scale'), metric = 'ROC', trControl = ctrl)
knn_preds = predict(knn_model, X_chem, type ='prob')
knn_roc = roc(response = any_damage, predictor = knn_preds[,1])
knn_auc = knn_roc$auc[1]
knn = list(model = knn_model, roc = knn_roc, auc = knn_auc)
```

```{r}
set.seed(100)
nb_model = train(X_chem, any_damage, method = 'nb', metric = 'ROC', trControl = ctrl)
nb_preds = predict(nb_model, X_chem, type = 'prob')
nb_roc = roc(response = any_damage, predictor = nb_preds[,1])
nb_auc = nb_roc$auc[1]
nb = list(model = nb_model, roc = nb_roc, auc = nb_auc)
```

```{r}
chem_result = list( mda = mda, nnet=nnet, svm=svm, knn=knn, nb=nb )

df = rbind(data.frame(name = 'MDA', auc = chem_result$mda$auc),
           data.frame(name = 'NNET', auc = chem_result$nnet$auc),
           data.frame(name = 'SVM', auc = chem_result$svm$auc),
           data.frame(name = 'KNN', auc = chem_result$knn$auc),
           data.frame(name = 'NB', auc = chem_result$nb$auc))
df
```

```{r}
varImp(chem_result$nnet$model, scale = F)
```

```{r}
plot(bio_result$mda$roc, legacy.axes = T, add = F, col = 'gray')
plot(chem_result$nnet$roc, legacy.axes = T, add =  T, col = 'darkgray')
```


##2 
Use fatty acid data from previous exercise set
a - use same data splitting and preprocessing. Which model has the best predictive ability? How does optimal model compare to linear models. would you infer that data have non linear separation
b - which oil type does optimal model accurately predict? least accurately predict?

```{r}
library(caret)
library(AppliedPredictiveModeling)
data(oil)
```

```{r}
zv_cols = nearZeroVar(fattyAcids)
findLinearCombos(fattyAcids)
```
## Build models

##mda
```{r}
set.seed(100)
mda_model = train(fattyAcids, oilType, method = 'mda', tuneGrid = expand.grid(.subclasses = 1:3))
mda_preds = predict(mda_model, fattyAcids)
mda_cm = confusionMatrix(data = mda_preds, reference = oilType)
mda = list(model = mda_model, predictions = mda_preds, confusion_matrix = mda_cm)
```

#NN
```{r}
nnet_grid = expand.grid(.size = 1:5, .decay = c(0, .1,1,2))
set.seed(100)
nnet_model = train(fattyAcids, oilType, method = 'nnet', tuneGrid = nnet_grid, preProcess =c("center","scale","spatialSign"),trace = F, maxit = 2000)
nnet_preds = predict(nnet_model, fattyAcids)
nnet_cm = confusionMatrix(data = nnet_preds, reference = oilType)
nnet = list(model = nnet_model, predictions = nnet_preds, confusion_matrix = nnet_cm)
```

##SVm
```{r}
library(kernlab)
sigma_est = sigest(as.matrix(fattyAcids))
svm_grid = expand.grid(.sigma = sigma_est[1], .C = 2^seq(-4.4))

set.seed(100)
svm_model = train(fattyAcids, oilType, method = 'svmRadial', tuneGrid = svm_grid, preProcess =c("center","scale"))
svm_preds = predict(svm_model, fattyAcids)
svm_cm = confusionMatrix(data = svm_preds, reference = oilType)
svm = list(model = svm_model, predictions = svm_preds, confusion_matrix = svm_cm)
```

#KNN
```{r}
set.seed(100)
knn_model = train(fattyAcids, oilType, method = 'knn', tuneLength = 30, preProcess =c("center","scale"))
knn_preds = predict(knn_model, fattyAcids)
knn_cm = confusionMatrix(data = knn_preds, reference = oilType)
knn = list(model = knn_model, predictions = knn_preds, confusion_matrix = knn_cm)
```
#Naive bayes
```{r}
nb_grid = expand.grid(fL = c(1,2), usekernel = c(T,F), adjust = T)
set.seed(100)
nb_model = train(fattyAcids, oilType, method = 'nb', tuneGrid = nb_grid)
nb_preds = predict(nb_model, fattyAcids)
nb_cm = confusionMatrix(data = nb_preds, reference = oilType)
nb = list(model = nb_model, predictions = nb_preds, confusion_matrix = nb_cm)
```
```{r}
oil_result = list(mda = mda, nnet = nnet, svm = svm, knn = knn)

df = rbind( data.frame(name="MDA", Accuracy=oil_result$mda$confusion_matrix$overall[1]),
            data.frame(name="NNET", Accuracy=oil_result$nnet$confusion_matrix$overall[1]),
            data.frame(name="SVM", Accuracy=oil_result$svm$confusion_matrix$overall[1]),
            data.frame(name="KNN", Accuracy=oil_result$knn$confusion_matrix$overall[1]))
df
```

```{r}
oil_result$nnet$confusion_matrix
```


















