---
title: "Chap2 - Short tour for predictive models"
author: "Me"
date: "8/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
library(AppliedPredictiveModeling)
data(FuelEconomy)
```


Format data for plotting against engine displacement

```{r}
library(dplyr)
library(tidyr)
#sort by engine displacement
cars2010 = cars2010[order(cars2010$EngDispl),]
cars2011 = cars2010[order(cars2010$EngDispl),]

#combine the data into one data frame.
cars2010a = cars2010 %>%
  mutate(Year = '2010 model year')

cars2011a = cars2011 %>%
  mutate(Year = '2011 model year')

#combine the data
carsdata = cars2010a %>%
  rbind(cars2011a)

#plot
library(lattice)
xyplot(FE ~ EngDispl|Year, carsdata, 
       xlab = 'Engine Displacement',
       ylab = 'Fuel Efficiency',
       between = list(x = 1.2))


```

```{r}
library(ggplot2)
ggplot(data = carsdata, aes(x = EngDispl, y = FE))+
  geom_point()+
  facet_wrap(~Year)
```

Fit a linear model and conduct 10 fold CV to estimate error

```{r}
library(caret)
set.seed(1)

lm1Fit = train(FE ~ EngDispl, data = cars2010, 
               method = 'lm', trControl = trainControl(method = 'cv'))
summary(lm1Fit)
```
```{r}
cars2010 %>%
  mutate(predFE = predict(lm1Fit,cars2010)) %>%
  ggplot()+
  geom_point(aes(EngDispl, FE))+
  geom_line(aes(EngDispl, predFE))

```

Fit a quadratic model
```{r}
#Create a sqaured variable of ED
cars2010 = cars2010 %>%
  mutate(ED2 = EngDispl^2)

cars2011 = cars2011 %>%
  mutate(ED2 = EngDispl^2)

#fit quadratic model
lm2Fit = train(FE ~ EngDispl + ED2, data = cars2010,
               method = 'lm', trControl = trainControl(method = 'cv'))
summary(lm2Fit)
```
The quadratic model produced lower RMSE

Let's plot the quadratic model

```{r}
cars2010 = cars2010 %>%
  mutate(q_pred = predict(lm2Fit, cars2010))

ggplot(cars2010)+
  geom_point(aes(EngDispl, FE))+
  geom_line(aes(EngDispl, q_pred))
```
Might over estimate the values on high EngDispl

Fit a MARS model 
```{r}
library(earth)
set.seed(1)
marsFit = train(FE ~ EngDispl,
                data = cars2010,
                method = 'earth',
                tuneLength = 15,
                trControl = trainControl(method = 'cv'))
summary(marsFit)
```

```{r}
plot(marsFit)
```

5 Terms has the lowest RMSE

Let's predict the test data for each model
```{r}
library(readr)
cars2011$lm1 = predict(lm1Fit, cars2011)
cars2011$lm2 = predict(lm2Fit, cars2011)
cars2011$mars = predict(marsFit, cars2011)
```


```{r}
cars2011 %>%
  gather(key = 'model', value = 'value', lm1, lm2, mars) %>%
  ggplot()+
  geom_point(aes(EngDispl, FE))+
  geom_line(aes(EngDispl, value, color = model))
```

Get test set performance value via caret's postResample function

```{r}
postResample(pred = cars2011$lm1, obs = cars2011$FE)
postResample(pred = cars2011$lm2, obs = cars2011$FE)
postResample(pred = cars2011$mars, obs = cars2011$FE)
```
Lowest error for Mars











